% main.tex — Vitrine v5: Capture-Adaptive Video-to-3D-Scene Pipeline for UE 5.8
% Academic report compiled with pdflatex + bibtex (4-pass).
%
% House style inherited from report/main_v4.tex:
%   documentclass report, a4paper, 11pt
%   Packages: inputenc, fontenc, babel, microtype, geometry, lmodern, inconsolata,
%             xcolor, graphicx, tikz, booktabs, tabularx, longtable, array,
%             amsmath, amssymb, csquotes, enumitem, caption, hyperref, bookmark,
%             fancyhdr, listings, tcolorbox, titlesec
%   Colour macros: UoSRed, DreamLabAccent, DarkBg, MidGrey, LightGrey, CodeBg,
%                  SuccessGreen, WarningAmber, DangerRed
%   Custom environments: keyfinding, technote (and recommendation from v4)
%   L{} column type
%   lstlisting bash configuration
%
% Dr John O'Hare, DreamLab AI Consulting Ltd — June 2026

\documentclass[a4paper,11pt]{report}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[british]{babel}
\usepackage{microtype}
\usepackage[a4paper,top=28mm,bottom=28mm,inner=26mm,outer=24mm,headheight=22pt]{geometry}
\usepackage{lmodern}
\usepackage[dvipsnames,svgnames,x11names]{xcolor}

\definecolor{UoSRed}{HTML}{B71234}
\definecolor{DreamLabAccent}{HTML}{6366F1}
\definecolor{DarkBg}{HTML}{0F172A}
\definecolor{MidGrey}{HTML}{64748B}
\definecolor{LightGrey}{HTML}{F1F5F9}
\definecolor{CodeBg}{HTML}{1E293B}
\definecolor{SuccessGreen}{HTML}{15803D}
\definecolor{WarningAmber}{HTML}{B45309}
\definecolor{DangerRed}{HTML}{B91C1C}

\usepackage{graphicx}
\graphicspath{{v5/}{figures/}{figures/v2/}{figures/v3/}}
\usepackage{tikz}
\usetikzlibrary{calc,positioning,arrows.meta,shapes.geometric,backgrounds,fit}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\usepackage{amsmath,amssymb}
\usepackage{csquotes}
\usepackage{enumitem}
\setlist{nosep,leftmargin=*}
\usepackage{caption}
\captionsetup{font=small,labelfont={bf,color=UoSRed}}

\usepackage[colorlinks=true,linkcolor=UoSRed,citecolor=DreamLabAccent,urlcolor=DreamLabAccent,
  bookmarks=true,bookmarksnumbered=true,
  pdfauthor={Dr John O'Hare},
  pdftitle={Vitrine: A Capture-Adaptive Video-to-3D-Scene Pipeline for Unreal Engine 5.8},
  pdfsubject={3D Gaussian Splatting, Unreal Engine 5.8, Capture-Adaptive Pipeline, Game Assets}]{hyperref}
\usepackage{bookmark}
\usepackage{url}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textcolor{MidGrey}{\leftmark}}
\fancyfoot[R]{\small\textcolor{MidGrey}{\thepage}}
\fancyfoot[C]{\small\textcolor{MidGrey}{Vitrine v5, DreamLab AI Consulting Ltd}}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.2pt}
\renewcommand{\headrule}{\hbox to\headwidth{\color{UoSRed}\leaders\hrule height \headrulewidth\hfill}}

\usepackage{listings}
\lstset{basicstyle=\ttfamily\footnotesize,backgroundcolor=\color{LightGrey},frame=single,
  rulecolor=\color{MidGrey},breaklines=true,showstringspaces=false,tabsize=2,
  keywordstyle=\color{DreamLabAccent}\bfseries,commentstyle=\color{MidGrey}\itshape,
  stringstyle=\color{SuccessGreen},extendedchars=true,
  literate={---}{{---}}3 {--}{{--}}2 {->}{{$\rightarrow$}}2
    {checkmark}{{\checkmark}}1 {>=}{{$\geq$}}2 {<=}{{$\leq$}}2}
\lstdefinestyle{bash}{language=bash,
  morekeywords={docker,curl,python3,git,cmake,cargo,pytest,docker-compose},
  keywordstyle=\color{DreamLabAccent}\bfseries}

\usepackage[most]{tcolorbox}
\newtcolorbox{keyfinding}[1][]{enhanced,colback=UoSRed!5,colframe=UoSRed,fonttitle=\bfseries,
  title={Key Finding},left=5mm,boxrule=0.8pt,#1}
\newtcolorbox{recommendation}[1][]{enhanced,colback=DreamLabAccent!6,colframe=DreamLabAccent,
  fonttitle=\bfseries,title={Recommendation},left=5mm,boxrule=0.8pt,#1}
\newtcolorbox{technote}[1][]{enhanced,colback=LightGrey,colframe=MidGrey,fonttitle=\bfseries,
  title={Technical Note},left=5mm,boxrule=0.6pt,#1}

\usepackage{titlesec}
\titleformat{\chapter}[display]{\normalfont\huge\bfseries\color{DarkBg}}
  {\textcolor{UoSRed}{\chaptertitlename\ \thechapter}}{8pt}{\Huge}
\titlespacing*{\chapter}{0pt}{-18pt}{26pt}
\titleformat{\section}{\normalfont\Large\bfseries\color{DarkBg}}{\textcolor{UoSRed}{\thesection}}{1em}{}
\titleformat{\subsection}{\normalfont\large\bfseries\color{DarkBg}}{\textcolor{DreamLabAccent}{\thesubsection}}{1em}{}

% Convenience macros
\newcommand{\chip}[2]{\textcolor{#1}{\textbf{#2}}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\validated}{\chip{SuccessGreen}{validated}}
\newcommand{\inprogress}{\chip{WarningAmber}{in progress}}
\newcommand{\designed}{\chip{MidGrey}{designed}}
\newcommand{\wip}{\chip{WarningAmber}{in progress}}
\newcommand{\built}{\chip{SuccessGreen}{built}}

% natbib-compatible bibliography style
\usepackage[numbers,sort&compress]{natbib}

\begin{document}
\sloppy

% ───────────────────────── Title page ─────────────────────────
\begin{titlepage}
\begin{tikzpicture}[remember picture, overlay]
  \fill[DarkBg] (current page.south west) rectangle (current page.north east);
  \foreach \r in {2,3.2,4.4,5.6} {
    \draw[UoSRed, opacity=0.15, line width=0.6pt]
      ($(current page.north west) + (2.5,-3)$) circle (\r);}
  \foreach \i in {0,0.8,...,6.4} {
    \draw[DreamLabAccent, opacity=0.10, line width=0.5pt]
      ($(current page.south east) + (-\i-1, 0.5)$) -- ($(current page.south east) + (-0.5, \i+1)$);}
  \fill[UoSRed] ($(current page.west) + (0,2)$) rectangle ($(current page.east) + (0,2.08)$);
  \fill[DreamLabAccent] ($(current page.west) + (0,1.9)$) rectangle ($(current page.east) + (0,1.94)$);
  \node[anchor=north west, text=UoSRed, font=\footnotesize\sffamily\bfseries, text width=13cm]
    at ($(current page.north west) + (3,-2)$)
    {v5 ACADEMIC REPORT \textcolor{MidGrey}{|} DREAMLAB REFERENCE RUN VALIDATED};
  \node[anchor=west, text=white, font=\fontsize{40}{44}\selectfont\bfseries\sffamily, text width=15cm]
    at ($(current page.west) + (3,5.6)$) {Vitrine};
  \node[anchor=west, text=white, font=\fontsize{13}{17}\selectfont\sffamily, text width=14cm]
    at ($(current page.west) + (3,3.2)$)
    {A Capture-Adaptive Video-to-3D-Scene Pipeline\\for Unreal Engine~5.8};
  \node[anchor=west, text=MidGrey, font=\normalsize\sffamily, text width=13cm]
    at ($(current page.west) + (3,0.5)$)
    {video \textbullet\ COLMAP \textbullet\ 3D Gaussian Splatting \textbullet\ SAM3 \textbullet\ TRELLIS.2\\
     NanoGS \textbullet\ mesh \textbullet\ xatlas bake \textbullet\ FBX \textbullet\ UE\,5.8 game assets};
  \node[anchor=west, text=white, font=\sffamily, text width=10cm]
    at ($(current page.west) + (3,-2.2)$)
    {\textbf{Dr John O'Hare}\\[2pt] DreamLab AI Consulting Ltd\\[8pt]
     \textcolor{MidGrey}{Pairs with the upstream}\\[2pt] \textbf{LichtFeld Studio} reconstruction engine};
  \node[anchor=west, text=DreamLabAccent, font=\sffamily\bfseries]
    at ($(current page.west) + (3,-5.0)$) {June 2026};
  \node[anchor=south east, text=MidGrey, font=\tiny\sffamily]
    at ($(current page.south east) + (-3,1)$)
    {v5.0 \textcolor{MidGrey}{|} CAPTURE-ADAPTIVE PIPELINE \textbullet\ UE\,5.8 DELIVERY};
\end{tikzpicture}
\end{titlepage}

% ───────────────────────── Front matter ─────────────────────────
\pagenumbering{roman}

% ── Abstract ──────────────────────────────────────────────────────────────────
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
\markboth{Abstract}{Abstract}

Vitrine is a \textbf{capture-adaptive} video-to-3D-scene pipeline that turns
hand-held walkthrough footage into structured, game-ready assets for
\textbf{Unreal Engine~5.8}.  Rather than applying a fixed reconstruction recipe
to every capture, Vitrine first \emph{diagnoses} each dataset on four
orthogonal axes --- sharpness, coverage, available hardware, and deliverable
format --- and then selects the combination of components whose benefit matches
the diagnosed bottleneck.  The key insight is that the best pipeline is a
function of the data, not a property of the system.

The pipeline is validated end-to-end on the \textbf{dreamlab reference capture}:
a 600-frame, 4K/30\,fps handheld walkthrough of a workshop/studio space,
reconstructed with COLMAP (600/600 frames registered) and trained to a
2.48M-Gaussian splat.  The validated dreamlab run produces: a real-Gaussian splat
room rendered in UE\,5.8 via the \textbf{NanoGS} plugin (MIT, recompiled for
UE\,5.8); three TRELLIS.2 textured object FBXs (chair, vacuum, toolbox) placed at
pre-computed real-world coordinates; and a comparative TSDF mesh branch.

A central empirical finding is \textbf{honest and negative}: the dreamlab room is
\emph{capture-limited}, not extractor-limited.  2DGS surface reconstruction
achieves planarity 69.7\% versus TSDF 72.0\% --- a $\Delta=2.3$ percentage-point
gap within measurement noise.  Neither exceeds the fair-room threshold; the binding
constraint is motion blur in the source footage (MUSIQ $\approx$31), and no mesh
extractor can raise that ceiling.  The room is therefore delivered as the
LichtFeld-native splat hero via NanoGS; the polygonal-mesh mandate is satisfied by
the TRELLIS.2 object FBXs.

Two further contributions are reported: a \textbf{directional sharpness gate}
(structure-tensor minor eigenvalue $\lambda_2$) that distinguishes directional
motion blur from isotropic softness and raises the bottom-quartile sharpness of
selected frames by 20--38\%; and a full hardware characterisation of
\textbf{ArtiFixer}~\cite{delutio2026_artifixer} on the NVIDIA RTX~6000~Ada
(48\,GB, sm\_89), showing that three targeted OOM fixes reduce peak VRAM from
out-of-memory to 45.5\,GB, enabling the 14B coverage-gap refiner to run on
Ada-generation hardware.

\tableofcontents
\listoffigures
\listoftables

% ───────────────────────── Main matter ─────────────────────────
\cleardoublepage
\pagenumbering{arabic}

% ====================================================================
\chapter{Introduction}
\label{chap:intro}

% ── Problem ──────────────────────────────────────────────────────────────────
\section{The Problem: Real Captures Fail in Different Ways}
\label{sec:intro:problem}

Photorealistic 3D reconstruction from hand-held video is a mature technology: the
combination of Structure-from-Motion (SfM), 3D Gaussian Splatting
(3DGS)~\cite{kerbl2023_3dgs}, and game-engine delivery is now within reach of any
operator with a modern smartphone and a GPU workstation.  The unsolved problem is
not the algorithm --- it is the \emph{mismatch between the algorithm's assumptions
and the real-world data it receives}.

Real captures fail in qualitatively different ways.  A handheld walkthrough at
walking pace fails due to \emph{motion blur}: the Gaussian field faithfully records
blurry observations, and no mesh extractor or training recipe can recover
sub-shutter-speed detail from pixels that were never resolved.  A tripod scan of a
large space fails due to \emph{coverage gaps}: the camera never reached certain
surfaces, Gaussians sprout as floaters in the unobserved void, and the mesh
contains holes.  A mobile-phone RGB-only capture fails differently from an
iPhone-LiDAR capture.  A project that demands Nanite-compatible collision meshes
fails differently from one that accepts a volumetric splat.

Any single fixed reconstruction recipe will be sub-optimal for most of these cases.
Optimising the pipeline for blur-limited footage (e.g.\ BAD-Gaussians
deblur~\cite{zhao2024_badgaussians}) adds wasted compute on a dense-but-sharp
capture.  Applying ArtiFixer~\cite{delutio2026_artifixer} to a blur-limited room is
similarly wasted: the model stays faithful to observed pixels and only fills
\emph{unobserved} regions; on a dense-but-blurry capture, there are no unobserved
regions to fill.

% ── Thesis ──────────────────────────────────────────────────────────────────
\section{The Capture-Adaptive Thesis}
\label{sec:intro:thesis}

Vitrine's answer is the \textbf{capture-adaptive thesis}: \emph{diagnose first,
then route}.  Every incoming capture is profiled on four independent axes before
any 3D computation begins:

\begin{description}
  \item[Sharpness] Is motion blur the dominant defect, or is the footage sharp?
    Measured by MUSIQ no-reference IQA~\cite{ke2021_musiq} and a full-resolution
    variance-of-Laplacian score via \texttt{pyiqa}~\cite{chen2022_pyiqa}.
  \item[Coverage] Are under-observed regions (holes, floaters) the dominant
    defect, or is the reconstruction dense?  Measured by a per-surface-point
    viewpoint-count occupancy probe.
  \item[Hardware] Does the capture include a depth sensor (iPhone LiDAR, event
    camera), or is it RGB-only?
  \item[Deliverable] Is a volumetric splat room representation sufficient, or does
    the project mandate a polygonal mesh (for collision, LOD, Nanite)?
\end{description}

\noindent A \emph{router} (Figure~\ref{fig:router}) evaluates these four axes
independently and assembles the appropriate configuration.  The four axes are
orthogonal: sharpness and coverage can each be good or bad independently, hardware
is a given, and the deliverable requirement is a project constraint.  The
router's decision space has $2^4 = 16$ nominal configurations; the section
files describe the validated subset.

% ── Contributions ─────────────────────────────────────────────────────────────
\section{Contributions}
\label{sec:intro:contributions}

This report documents the following validated contributions:

\begin{enumerate}
  \item \textbf{Capture-adaptive router} (\S\ref{sec:method:overview}).  A
    four-axis diagnosis-and-route mechanism that selects pipeline components based
    on the capture profile rather than a fixed recipe.

  \item \textbf{NanoGS splat-hero room in UE~5.8}
    (\S\ref{sec:results:nanogs}).  The LichtFeld-native 3DGS output is cleaned with
    SuperSplat and rendered as real Gaussians inside UE~5.8 via the NanoGS plugin,
    recompiled from its UE~5.6/5.7 codebase.  This closes the prior ``LiDAR
    point-cloud workaround'' and delivers the highest fidelity available from the
    captured data without an intermediate mesh-extraction step.

  \item \textbf{TRELLIS.2 object FBXs in UE~5.8}
    (\S\ref{sec:results:objects}).  Three dreamlab objects (chair, vacuum,
    toolbox) are reconstructed as textured polygonal game assets via the
    TRELLIS.2~\cite{xiang2025_trellis2} \texttt{hull\_e2e} workflow and placed at
    pre-computed real-world coordinates, satisfying the Nanite polygonal-mesh
    mandate without requiring the room to be meshed.

  \item \textbf{Empirical capture-limited finding}
    (\S\ref{sec:results:mesh}).  A controlled comparison of 2DGS and TSDF
    mesh extraction on the dreamlab capture establishes that the room is
    capture-limited: SOTA surface reconstruction provides no measurable benefit
    over a legacy TSDF on motion-blurred input.

  \item \textbf{Directional sharpness gate}
    (\S\ref{sec:results:framegate}).  An upgraded frame-quality gate based on the
    structure-tensor minor eigenvalue $\lambda_2$ that distinguishes
    directional motion blur from isotropic softness and raises the bottom-quartile
    sharpness of selected frames by 20--38\%.

  \item \textbf{ArtiFixer on Ada sm\_89}
    (\S\ref{sec:results:artifixer}).  Three targeted VRAM fixes reduce
    ArtiFixer~\cite{delutio2026_artifixer} peak usage from OOM to 45.5\,GB on a
    48\,GB RTX~6000~Ada, enabling the coverage-gap refiner to run on Ada-generation
    hardware without a FlashAttention-3 dependency.
\end{enumerate}

\begin{keyfinding}
The dreamlab reference run validates the capture-adaptive architecture on real data.
The headline lesson is honest: a SOTA mesh extractor cannot raise the quality ceiling
set by the source footage.  The splat-hero path (NanoGS) is both the highest-fidelity
and the most pipeline-lean option for a blur-limited capture; recapture at shorter
shutter speed is the real quality lever.
\end{keyfinding}

% ====================================================================
\chapter{Technologies and Related Work}
\label{chap:tech}

\input{v5/sec_technologies}

% ====================================================================
\chapter{Method: The Capture-Adaptive Pipeline}
\label{chap:method}

\input{v5/sec_method}

% ====================================================================
\chapter{Results: Dreamlab Reference Run}
\label{chap:results}

\input{v5/sec_results}

% ====================================================================
% Figures — float here, referenced from chapters above
% ====================================================================

\input{v5/figures}

% ====================================================================
\appendix
\input{v5/appendix_playbook}

% ====================================================================
\bibliographystyle{unsrt}
\bibliography{v5/references}

\end{document}
